---
title: "Varholak_week6"
output: html_document
date: "2024-07-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``` {r loading_libraries, warning=FALSE}
# load the libraries for ggplot2, cowplot, and palmerpenguins
library(ggplot2)
library(palmerpenguins)
library(cowplot)
#make the penguin data into an object 
peng_df <- penguins
```

## Question 1 

1A. Violin plot with sex on X axis and flipper length on the Y axis and have the sexes be different colors. 

```{r hw_6_q1A, warning=FALSE}
#remove the penguins that do not have their sex recorded. 
peng_df2 <- dplyr::filter(.data = peng_df, sex == 'male' | sex == 'female')
#create a graph with peng_df where x is sex and y is flipper length. Change fill color for sex
ggplot(data = peng_df2, mapping = aes(x =sex, y =flipper_length_mm, fill = sex)) + 
  geom_violin()
```


1B. Divide the plot into three panels, one per species. Remove the gray background, keep the axis lines, and add axis names that remove the underscore from the raw data titles. 

```{r hw_6_q1B}
plot1 <- ggplot(data = peng_df2, mapping = aes(x =sex, y =flipper_length_mm, fill = sex)) + 
  geom_violin()+ 
  #Divide the plot into three panels (one per species) 
  facet_grid(. ~ species)+
  #remove the gray background, keep the axis lines. 
  theme(panel.grid = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(color = 'black'))+
  #add axis names that remove the underscore from raw data titles
  labs(x="Sex", y="Flipper Length (mm)")
plot1
```

## Question 2 

2A. Violin plot with flipper length on the x axis and island on the y axis. Fill color of each violin different between islands. Remove gray background, keep the black axis lines, and make the axis titles match the first graph. 

```{r hw_6_q2A, warning=FALSE}
#create a violin plot with flipper length on the x axis and island on the y axis where islands have different fill colors. 
ggplot(data = peng_df, mapping = aes(x = flipper_length_mm, y =island, fill = island)) + 
  geom_violin()+ 
  #remove the gray background, keep the axis lines
  theme(panel.grid = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(color = 'black'))+
  #add axis names that remove the underscore from raw data titles
  labs(x= 'Flipper Length (mm)', y='Island')
```

2B. Pick a custom color for each of the three islands and manually set the violin fill to give each island a custom color. 

```{r hw_6_q2B, warning=FALSE}
#custom color per island
plot2 <- ggplot(data = peng_df, mapping = aes(x = flipper_length_mm, y =island, fill = island))+
  geom_violin()+ 
  theme(panel.grid = element_blank(), 
        panel.background = element_blank(), 
        axis.line = element_line(color = 'black'))+
  #add axis names that remove the underscore from raw data titles
  labs(x= 'Flipper Length (mm)', y='Island')+
  #customize the colors so they are different from the default colors. 
  scale_fill_manual(values = c('Biscoe' = '#a0c4ff', 'Dream' = '#ffc6ff', 'Torgersen' = '#fdffb6'))
plot2
```

3. Combine the plots from 1B and 2B into one graph. 

```{r hw_6_q3, warning=FALSE}
#combine two of the previous plots into one graph. 
cowplot::plot_grid(plot1, plot2, ncol = 1, nrow = 2)
```





