---
title: "Varholak_week7"
output: html_document
date: "2024-07-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

1A. Make an object containing the number 35 and convert it to Celsius (C)
```{r hw_7_q1A}
#Make an object with the number 35. 
object_f <- 35
#Convert to celsius 
object_c <-  (object_f - 32) * 5/9
object_c
```

1B. Make a for loop to convert a vector of numbers from Fahrenheit(F) to Celsius (C). 

``` {r hw_7_q1B}
numbers_vec <- c(35, 45, 55, 65, 75, 85, 95)
#for loop to convert vector values from F to C
for (i in numbers_vec) {
  object_c1 <- (i - 32) * 5/9
  print(object_c1)
}
  
```

1C. Add a conditional where if the temperature in Celsius is less than or equal to 18 a message prints that includes the temperature in Fahrenheit and a note recommending bringing a jacket. 

``` {r hw_7_q1C}
numbers_vec <- c(35, 45, 55, 65, 75, 85, 95)
#for loop to convert vector values from F to C with conditionals in if statments
for (i in numbers_vec) {
  object_c1 <- (i - 32) * 5/9
  
  if (object_c1 < 18 | object_c1 == 18) {
    print (paste(object_c1, "degrees Celsius is", i, "degrees Fahrenheit, you should bring a jacket"))
  }else  {
      print(paste(object_c1, "degrees Celsius is", i, "degrees Fahrenheit, you do not need to bring a jacket"))
  }
}
```

1D. If the degrees (C) is less than or equal to 18, write a message about bringing a jacket. If the degrees (C) is in between 18 and 23 write a message saying that neither shorts nor a jacket are required. If degrees (C) is greater than or equal to 23, write a message recommending bringing shorts. 
```{r hw_7_q1D}
numbers_vec <- c(35, 45, 55, 65, 75, 85, 95)
#for loop to convert vector values from F to C with an additional conditional using else if.
for (i in numbers_vec) {
  object_c1 <- (i - 32) * 5/9
  
  if (object_c1 < 18 | object_c1 == 18) {
    print (paste(object_c1, "degrees Celsius is", i, "degrees Fahrenheit, you should bring a jacket"))}
  else if (object_c1 > 18 & object_c1<23) {
    print (paste(object_c1, "degrees Celsius is", i, "degrees Fahrenheit, neither shorts nor a jacket are needed"))
  }
  else  {
      print(paste(object_c1, "degrees Celsius is", i, "degrees Fahrenheit, you should pack shorts"))
  }
}
```


